ChangeScores

### 2023-05-30 DataSetup.Rmd
Cleaned up file organization

#t test for age differences
t.test(sample_1_baseline$interview_age, sample_2_baseline$interview_age)

@@

#t test for age differences
t.test(sample_1_followup$interview_age, sample_2_followup$interview_age)

>>>>

#t test for age differences
t_apa(t_test(sample_1_baseline$interview_age, sample_2_baseline$interview_age))

#chi-sq for sex differences
#chisq_apa(chisq.test(sample_1_baseline$sex, sample_2_baseline$sex))

prop <- c(length(which(sample_2_baseline$sex == "F")) / length(which(sample_2_baseline$sex == "F" | sample_2_baseline$sex == "M")), length(which(sample_2_baseline$sex == "M")) / length(which(sample_2_baseline$sex == "F" | sample_2_baseline$sex == "M")))

chisq_apa(chisq.test(x = table(sample_1_baseline$sex), p = prop))

@@

#t test for age differences
t_apa(t_test(sample_1_followup$interview_age, sample_2_followup$interview_age))

#chi-sq for sex differences

props <- c(length(which(sample_2_followup$sex == "F")) / length(which(sample_2_followup$sex == "F" | sample_2_followup$sex == "M")), length(which(sample_2_followup$sex == "M")) / length(which(sample_2_followup$sex == "F" | sample_2_followup$sex == "M")))

chisq_apa(chisq.test(x = table(sample_1_followup$sex), p = props))

### 2023-05-30 Difference Scores /ABCDDiffScoreAnalyses.Rmd
Cleaned up file organization

load("~/Volumes/DevBrainLab/Lucy/fyp/ABCDBrainAgeDiffCorrected.Rda")
#brain_age_corrected_df_diff

@@

diff_pred_plot <- ggplot(brain_age_corrected_df_diff, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#619B8A")+
  geom_abline(a=0, b=1, color='#233D4D') +
  labs(x = "Scan/Chronological Age (Midpoint)", y ="Predicted Age (Midpoint)") +
  theme_classic() + 
  theme(text = element_text(size = 14))  
#ggsave(filename="diff_truth_corrected_pred_mod1.png",
  #                        width=7, height=4, units='in', dpi=300)

>>>>

load("/Volumes/devbrainlab/Lucy/BrainAGE/Diff Scores/ABCDBrainAgeDiffCorrected.Rda")
#brain_age_corrected_df_diff

@@

diff_pred_plot <- ggplot(brain_age_corrected_df_diff, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#619B8A")+
 # geom_abline(a=0, b=1, color='#233D4D') +
  labs(x = "Scan/Chronological Age (Midpoint)", y ="Predicted Age (Midpoint)") +
  theme_classic() + 
  theme(text = element_text(size = 14))  
#ggsave(filename="diff_truth_corrected_pred_mod1.png",
  #                        width=7, height=4, units='in', dpi=300)

diff_pred_plot

cor(brain_age_corrected_df_diff$truth, brain_age_corrected_df_diff$corrected_pred)

# uncorrected 
ggplot(brain_age_corrected_df_diff, aes(x=truth, y=.pred)) + 
  geom_jitter(color="#619B8A")+
 # geom_abline(a=0, b=1, color='#233D4D') +
  labs(x = "Scan/Chronological Age (Midpoint)", y ="Predicted Age (Midpoint)") +
  theme_classic() + 
  theme(text = element_text(size = 14))

### 2023-05-30 Difference Scores /DifferenceScores.Rmd
Cleaned up file organization

Prep for training
```{r}
# Select only difference columns
smri_difference_calculated_subset <- smri_difference_calculated %>% 
  select(src_subject_id, ends_with("diff")) 
  #divide all brain columns by interview age

smri_difference_calculated_subset[, -(1:2)] <- sweep(smri_difference_calculated_subset[, -(1:2)], 1, smri_difference_calculated_subset[, 2], "/")

@@

>>>>

Prep for training
```{r}
# Select only difference columns
smri_difference_calculated_subset <- smri_difference_calculated %>% 
  select(src_subject_id, ends_with("diff"))

#divide all brain columns by change in interview age
smri_difference_calculated_subset[, -(1:2)] <- sweep(smri_difference_calculated_subset[, -(1:2)], 1, smri_difference_calculated_subset[, 2], "/")

@@

############################################################
MAY 2023 Update - New Approach
###########################################################

Updates to this version: Predicting T2 age from difference scores (scaled by time interval, in delta interview age)

Load Structural Data
```{r}
#Load list of included subjects
struc_include_baseline<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:11) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include)) %>% 
  filter(imgincl_t1w_include == 1) %>% 
  filter(eventname=="baseline_year_1_arm_1")

struc_include_followup<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:11) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include)) %>% 
  filter(imgincl_t1w_include == 1) %>% 
  filter(eventname=="2_year_follow_up_y_arm_1")

#Load structural data
#Filter for volume & area measurements
smri_baseline_v2 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
  filter(eventname=="baseline_year_1_arm_1") %>% 
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include_baseline$src_subject_id) %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area|thick")) %>% 
  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))

smri_followup_v2 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
  filter(eventname=="2_year_follow_up_y_arm_1") %>% 
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include_followup$src_subject_id) %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area|thick")) %>% 
  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))

#renamed_smri_baseline<-rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/renamed_smri_baseline.Rda")

Load and join data
```{r}

#Convert to numeric
smri_baseline_v2[,6:ncol(smri_baseline_v2)] <- sapply(smri_baseline_v2[,6:ncol(smri_baseline_v2)],as.numeric)

# Calculate age in years
smri_baseline_v2 <- smri_baseline_v2 %>% 
  mutate(interview_age = as.numeric(interview_age)/12)

#Convert to numeric
smri_followup_v2[,6:ncol(smri_followup_v2)] <- sapply(smri_followup_v2[,6:ncol(smri_followup_v2)],as.numeric)

# calculate age in years
smri_followup_v2 <- smri_followup_v2 %>% 
  mutate(interview_age = as.numeric(interview_age)/12)

smri_difference_v2 <- left_join(smri_baseline_v2, smri_followup_v2, by ="src_subject_id")
```

Calculate difference scores
```{r}
smri_difference_calculated_v2 <- smri_difference_v2 %>%
 select(-c(subjectkey.x,subjectkey.y, sex.x,sex.y, eventname.x,eventname.y)) %>% 
 mutate(across(ends_with(".y"), .names = "{col}_diff") - across(ends_with(".x"))) %>%
 rename_with(~ sub("_\\d+", "", .), ends_with("_diff"))

Set up data
```{r}
smri_difference_v2 <- smri_difference_calculated_v2 %>% 
  select(src_subject_id, interview_age.y, ends_with("diff"))

#divide all brain columns by change in interview age
# this gets us per year change (not a %, but numeric change per year)
smri_difference_v2[, -(1:3)] <- sweep(smri_difference_v2[, -(1:3)], 1, smri_difference_v2[, 3], "/")

smri_difference_v2 <- smri_difference_v2 %>% 
  filter(!is.na(interview_age.y_diff))

#Split sample
smp_size_diff_v2 <- floor(0.25 * nrow(smri_difference_v2)) #training sample of 1000 participants

## set the seed to make your partition reproducible
set.seed(123)
sample_ind_diff_v2 <- sample(seq_len(nrow(smri_difference_v2)), size = smp_size_diff_v2)

sample_train_diff_v2 <- smri_difference_v2[sample_ind_diff_v2, ]
sample_test_diff_v2 <- smri_difference_v2[-sample_ind_diff_v2, ]

# select just subject_id, brain columns, age midpoint
sample_train_diff_v2 <- sample_train_diff_v2 %>% 
select(-c(src_subject_id, interview_age.y_diff))

save(sample_train_diff_v2, file="sample_train_diff_v2.Rda")
save(sample_test_diff_v2, file="sample_test_diff_v2.Rda")
```
